# Silvicultural Metrics{#silv}

Overstory metrics are calculated using the UAS SfM-derived DBH and height measurements, where DBH is estimated using a local DBH-Height allometric relationship. In the following analysis and reporting, overstory includes any tree where the DBH is `r scales::comma(ostory_dbh_cm/2.54,accuracy = 0.1)` in. (`r scales::comma(ostory_dbh_cm,accuracy = 0.01)` cm) or greater. Understory includes any tree where the DBH is less than `r scales::comma(ostory_dbh_cm/2.54,accuracy = 0.1)` in. (`r scales::comma(ostory_dbh_cm,accuracy = 0.01)` cm) or greater. 

```{r silv-calc, results='hide'}
# Common silvicultural metrics are calculated for the entire extent.
  # Note, that stand-level summaries can be computed if stand vector data is provided.
  # metrics include:
    # "n_trees"
    # "stand_area_ha"
    # "trees_per_ha"
    # "mean_dbh_cm"
    # "qmd_cm"
    # "mean_tree_height_m"
    # "loreys_height_m"
    # "basal_area_m2"
    # "basal_area_m2_per_ha"
# filter trees based on stand area
harvests_trees = stand_sf %>% 
  dplyr::mutate(
    stand_area_m2 = sf::st_area(.) %>% as.numeric()
    , stand_area_ha = stand_area_m2/10000
    , suid = "1"
  ) %>% 
  dplyr::select(
    suid, stand_area_m2, stand_area_ha
  ) %>% 
  sf::st_intersection(treetops_sf_with_dbh)

### stand-level summaries
  silv_metrics = harvests_trees %>%
    sf::st_drop_geometry() %>%
    # dplyr::filter(dbh_cm >= ostory_dbh_cm) %>% 
    dplyr::ungroup() %>%
    dplyr::group_by(suid,stand_area_ha) %>%
    dplyr::summarise(
      n_trees = dplyr::n_distinct(treeID)
      , mean_dbh_cm = mean(dbh_cm, na.rm = T)
      , mean_tree_height_m = mean(tree_height_m, na.rm = T)
      , loreys_height_m = sum(basal_area_m2*tree_height_m, na.rm = T) / sum(basal_area_m2, na.rm = T)
      , basal_area_m2 = sum(basal_area_m2, na.rm = T)
      , sum_dbh_cm_sq = sum(dbh_cm^2, na.rm = T)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      trees_per_ha = (n_trees/stand_area_ha)
      , basal_area_m2_per_ha = (basal_area_m2/stand_area_ha)
      , qmd_cm = sqrt(sum_dbh_cm_sq/n_trees)
    ) %>%
    dplyr::select(-c(sum_dbh_cm_sq)) %>%
    # convert to imperial units
    calc_imperial_units_fn() %>%
    dplyr::select(
      "suid"
      , "n_trees"
      , "stand_area_ha"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2"
      , "basal_area_m2_per_ha"
      # imperial
      , "stand_area_ac"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2"
      , "basal_area_ft2_per_ac"
    )

### export tabular
  write.csv(
      silv_metrics
      , paste0(delivery_dir, "/stand_silv_metrics.csv")
      , row.names = F
    )

```

```{r}
tab_met_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ha"
      , "n_trees"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2_per_ha"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ha"
            , "n_trees"
            , "trees_per_ha"
            , "mean_tree_height_m"
            , "loreys_height_m"
            , "mean_dbh_cm"
            , "qmd_cm"
            , "basal_area_m2_per_ha"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ha)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ha^{-1}$")
            , latex2exp::TeX("Mean Tree Height (m)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (m)", output = "character")
            , latex2exp::TeX("Mean DBH (cm)", output = "character")
            , latex2exp::TeX("QMD (cm)", output = "character")
            , latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(
          parse = T, fill = "white", linecolor = "gray"
          , size = 10
        )
      )
    ) %>% 
    ggpubr::tab_add_title(
      text = paste0(
        "Silvicultural Metrics (metric)"
        # , "\nOverstory (>= "
        # , scales::comma(ostory_dbh_cm,accuracy = 0.1)
        # , " cm DBH)"
        )
      , face = "bold"
      , size = 10, padding = unit(0.5, "line")
    )
}
# tab_met_fn(1)
```

```{r}
tab_imp_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ac"
      , "n_trees"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2_per_ac"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ac"
            , "n_trees"
            , "trees_per_ac"
            , "mean_tree_height_ft"
            , "loreys_height_ft"
            , "mean_dbh_in"
            , "qmd_in"
            , "basal_area_ft2_per_ac"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ac)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ac^{-1}$")
            , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
            , latex2exp::TeX("Mean DBH (in)", output = "character")
            , latex2exp::TeX("QMD (in)", output = "character")
            , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(
          parse = T, fill = "white", linecolor = "gray"
          , size = 10
        )
      )
    ) %>% 
    ggpubr::tab_add_title(
      text = paste0(
        "Silvicultural Metrics (imperial)"
        # , "\nOverstory (>= "
        # , scales::comma(ostory_dbh_cm/2.54,accuracy = 0.1)
        # , " in. DBH)"
        )
      , face = "bold"
      , size = 10, padding = unit(0.5, "line")
    )
}
# tab_imp_fn(1)
```

## Stand Summary Metrics

```{r}
tab_imp_fn(1) + tab_met_fn(1)
# save it
ggplot2::ggsave(paste0(delivery_dir, "/silv_metrics.jpeg"), dpi = "print", height = 3.5, width = 6, device = "jpeg")
```

## DBH Distrubtion

```{r os-dbh-dist}
dbh_temp = harvests_trees %>%
  sf::st_drop_geometry() %>%
  dplyr::select(treeID, dbh_in) %>% 
  dplyr::ungroup() %>%
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = dbh_in)
  ) +
    geom_density(alpha = 0.8, fill = "navy", color = NA) + 
    labs(
      x = "DBH (in)"
      , y = ""
      , title = "SfM-derived tree DBH distribution"
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
    scale_y_continuous(breaks = NULL) +
    theme_light() +
    theme(
      legend.position = "none"
    )
dbh_temp
```

```{r os-dbh-dist2}
tpa_temp = 
  harvests_trees %>%
    sf::st_drop_geometry() %>%
    dplyr::select(treeID, dbh_in, stand_area_ha) %>% 
    dplyr::ungroup() %>%
    dplyr::distinct() %>% 
    dplyr::mutate(
      dbh_cl = ggplot2::cut_width(dbh_in, width = 1, center = 0.5, closed = "left")
    ) %>% 
    dplyr::group_by(dbh_cl) %>% 
    dplyr::summarise(
      n_trees = dplyr::n()
      , trees_per_ac = dplyr::n()/max(stand_area_ha %>% `*`(2.471))
      , trees_per_ha = dplyr::n()/max(stand_area_ha)
    ) %>%
    dplyr::ungroup() %>% 
    dplyr::mutate(
      tot_trees_per_ac = sum(trees_per_ac)
      , tot_trees_per_ha = sum(trees_per_ha)
      , pct = trees_per_ac/tot_trees_per_ac
      , dbh_cl_lab = paste0(
        stringr::word(dbh_cl, 1, sep = fixed(",")) %>% readr::parse_number()
        , " to "
        , stringr::word(dbh_cl, -1, sep = fixed(",")) %>% readr::parse_number() %>% `-`(0.1)
      ) %>% 
      factor() %>% 
      forcats::fct_reorder(stringr::word(dbh_cl, 1, sep = fixed(",")) %>% readr::parse_number())
    )

# plot
tpa_temp %>% 
  ggplot(
    mapping = aes(
      x = dbh_cl_lab, y = trees_per_ac
      , fill=trees_per_ac
      , label = paste0(
        scales::comma(trees_per_ac, accuracy = 0.1)
        , "\n"
        , scales::percent(pct, accuracy = 0.1)
      )
    )
  ) +
    geom_col(width = 0.7) +
    geom_text(color = "black", size = 3, vjust = -0.2) +
    scale_fill_viridis_c(option = "mako", direction = -1) +
    scale_y_continuous(labels = scales::comma_format(accuracy = 1),expand = expansion(mult = c(0, .13))) +
    labs(
      x = "DBH (in) Class"
      , y = "TPA"
      , title = "Trees Per Acre by Diameter Class"
      , subtitle = paste0(
          "Total TPA: "
          , scales::comma(tpa_temp$tot_trees_per_ac[1], accuracy = 1)
        )
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.text.x = element_text(angle = 90, size = 10, face = "bold", vjust = 0.5, hjust = 1)
    )
# save it
ggplot2::ggsave(paste0(delivery_dir, "/tpa_dbh.jpeg"), dpi = "print", height = 7, width = 10.5, device = "jpeg")
```

## Height Distribution

```{r os-ht-dist}
ht_temp = harvests_trees %>%
  sf::st_drop_geometry() %>%
  dplyr::select(treeID, tree_height_ft) %>% 
  dplyr::ungroup() %>%
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = tree_height_ft)
  ) +
    geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
    labs(
      x = "Height (ft)"
      , y = ""
      , title = "SfM-derived tree height distribution"
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
    scale_y_continuous(breaks = NULL) +
    theme_light() +
    theme(
      legend.position = "none"
    )
ht_temp
```

height + DBH

```{r}
ht_temp + dbh_temp
# save it
ggplot2::ggsave(paste0(delivery_dir, "/ht_dbh_dist.jpeg"), dpi = "print", height = 7, width = 10.5, device = "jpeg")
```

```{r}
ggplot(data = harvests_trees, mapping = aes(y = dbh_in, x = tree_height_ft)) +
  geom_point(size = 3, alpha = 0.8, color = "firebrick") + 
  scale_y_continuous(limits = c(0,NA), breaks = scales::extended_breaks(6)) +
  scale_x_continuous(limits = c(0,NA), breaks = scales::extended_breaks(15)) +
  labs(
    title = "SfM-derived tree DBH modeled by tree height"
    , x = "Height (ft)"
    , y = "DBH (in)"
  ) +
  theme_light()
# save it
ggplot2::ggsave(paste0(delivery_dir, "/ht_dbh_relationship.jpeg"), dpi = "print", height = 7, width = 10.5, device = "jpeg")
```


```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

